---
title: "STA-6257-PCA"
author: "Tavish Bryan"
format: html
editor: visual
---

## Method

Firstly, when performing PCA a new set of orthogonal coordinate axes are identified from the original data set, which is accomplished by finding the direction of maximal variance through each dimension. This is equivalent to using the least squares method to find the line of best fit. This new axis is the first principal component of the data set. Next we use orthogonal projection to project the coordinates onto the new axis. Once this is done we obtain a second principal component (and principal coordinate axis) by finding the direction of the second largest variance in the data, this axis is orthogonal to our first PC. These two PCs define a plane onto which we can project further coordinates onto (Richardson).

```{r}
install.packages("devtools")
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
library(AER)
library(tidyverse)
library(stats)
```

```{r}
data(Affairs)

pc = princomp(Affairs[,c(1,3:4,6:9)])
affairs.pca = prcomp(Affairs[,c(1,3:4,6:9)], 
                     center = TRUE, scale. = TRUE, rank. = 2)
summary(affairs.pca)
summary(pc)

str(affairs.pca)

ggbiplot(affairs.pca)
ggbiplot(pc)
#plot(pc)
#plot(affairs.pca)

autoplot(pc1,
  # AUTOPLOT OPTIONS
  data=conspiracy, 
  colour="PoliticalIdeology", 
  loadings=T, loadings.colour=alpha("#191970", 0.5), 
  loadings.label=T, loadings.label.colour="#191970", loadings.label.size=5, loadings.label.hjust=0) + 
  # CUSTOM COLORS FOR POLITICAL IDEOLOGY GROUPS
  scale_colour_manual(values = alpha(paletteer_d("rcartocolor::Temps"), 0.5)) +
  # GGPLOT THEME OPTIONS
  theme_bw() +
  theme(legend.key = element_rect(fill = "#FAFAF5"),
        legend.background = element_rect(fill = "#FAFAF5"),
        legend.text = element_text(color = "#8B814C", size = 14),
        legend.title = element_text(color = "#8B814C", size = 16),
        panel.border = element_rect(color = "#8B814C"),
        plot.background = element_rect(fill = "#FAFAF5"),
        axis.text = element_text(color = "#8B814C", size = 14),
        axis.title = element_text(color = "#8B814C", size = 16),
        axis.ticks = element_line(color = "#8B814C"))

```
